The Concurrent Calculi Formalisation Benchmark

### 2024-02-19 short-paper/short-paper.tex
Brit-ify spelling.

The POPLMark challenge~\cite{POPLMark}
played a pivotal role in advancing the field of proof assistants,
libraries, and best practices for programming language
mechanisation. By providing a shared framework for systematically
evaluating mechanisation techniques, it catalyzed a significant
shift towards publications that include mechanised proofs within the programming language research community.

@@

Our benchmark problems (detailed in \cref{app:challenges}) center around three key aspects that frequently pose challenges when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
Mechanisations must often address several of these
aspects at the same time; however, we see the combination of
techniques as a next step beyond the scope of this challenge, as
discussed further in \Cref{sec:going-beyond}.
Our benchmark challenges are based on process calculi, as these provide a simple but
realistic showcase of common issues encountered during mechanisation efforts.
While the three highlighted aspects are not the sole sources of difficulties, they emerge in most mechanisations of concurrency theory.

@@

\todo{Refer to the future benchmark section about other things one could do, but they still all need the basics}
Still, these (and other) applications need the basic
techniques that our challenge problems emphasize:
Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory. These problems center around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavors.
To achieve design goal \ref{item:goal-comperison-accessibility}, we have
formulated our challenge problems to explore the three aspects
(linearity, scope extrusion, coinduction) independently, so that they
may be solved individually and in any order; each problem is
reasonably small and easily understandable with basic knowledge of
textbook concurrency theory, process calculi, and type theory.  The
process calculus used in each challenge focuses on the features that
we want to emphasize, and omits all constructs (such as choices) that
would complicate the mechanisation without bringing additional
tangible benefits and insights.  The minimality and uniformity of the
calculi also allows us to target design goal
\ref{item:goal-tutorials}.

@@

In fact, coinduction is widely supported in proof assistants.
Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
Newer entrants to the field like the Lean proof assistant
are working on adding a notion of coinduction to their system~\cite{Avigad2019}%,Keizer2023}.
%
As with most features, proof assistants support
subtly different formalisms that enable different techniques and motivate this challenge.
Some systems provide more than one abstraction to use coinduction.
A good example of this are the parameterized coinduction~\cite{Hur2013}
and interaction tree~\cite{Xia2019} libraries for Coq.

>>>>

The POPLMark challenge~\cite{POPLMark}
played a pivotal role in advancing the field of proof assistants,
libraries, and best practices for programming language
mechanisation. By providing a shared framework for systematically
evaluating mechanisation techniques, it catalysed a significant
shift towards publications that include mechanised proofs within the programming language research community.

@@

Our benchmark problems (detailed in \cref{app:challenges}) centre around three key aspects that frequently pose challenges when mechanising concurrency theory: \emph{linearity}, \emph{scope extrusion}, and \emph{coinductive reasoning}.
Mechanisations must often address several of these
aspects at the same time; however, we see the combination of
techniques as a next step beyond the scope of this challenge, as
discussed further in \Cref{sec:going-beyond}.
Our benchmark challenges are based on process calculi, as these provide a simple but
realistic showcase of common issues encountered during mechanisation efforts.
While the three highlighted aspects are not the sole sources of difficulties, they emerge in most mechanisations of concurrency theory.

@@

\todo{Refer to the future benchmark section about other things one could do, but they still all need the basics}
Still, these (and other) applications need the basic
techniques that our challenge problems emphasise:
Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory. These problems centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavours.
To achieve design goal \ref{item:goal-comperison-accessibility}, we have
formulated our challenge problems to explore the three aspects
(linearity, scope extrusion, coinduction) independently, so that they
may be solved individually and in any order; each problem is
reasonably small and easily understandable with basic knowledge of
textbook concurrency theory, process calculi, and type theory.  The
process calculus used in each challenge focuses on the features that
we want to emphasise, and omits all constructs (such as choices) that
would complicate the mechanisation without bringing additional
tangible benefits and insights.  The minimality and uniformity of the
calculi also allows us to target design goal
\ref{item:goal-tutorials}.

@@

In fact, coinduction is widely supported in proof assistants.
Isabelle, Coq, and Abella seamlessly support coinductive definitions and reasoning.
Agda and Beluga support it through the convenient abstraction of co-patterns~\cite{Abel2013}.
Newer entrants to the field like the Lean proof assistant
are working on adding a notion of coinduction to their system~\cite{Avigad2019}%,Keizer2023}.
%
As with most features, proof assistants support
subtly different formalisms that enable different techniques and motivate this challenge.
Some systems provide more than one abstraction to use coinduction.
A good example of this are the parameterised coinduction~\cite{Hur2013}
and interaction tree~\cite{Xia2019} libraries for Coq.
